Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-05-09 VegOzone.tex
Second revision of Sec 3.2.2

Similarly, we find a skew towards higher \chem{[O_3]} for Svanvik in 2018, while 2019 follows a normal distribution rather well (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}). In conjunction with the meteorological conditions, we therefore can assume that 2019 had been a normal year and the annual average anomally in 2019 can be used for bias correction of the Svanvik climatology (\chem{\left<\Delta[O_3]\right>} = $(1.2\pm 5.0)\,\unit{ppb}$). To further strengthen this, we look at trends in ozone observations at Esrange and Pallas and compute the average difference between recent years and the mid 1990s (\chem{\Delta_{present-past}\left<[O_3]\right>}). Present is defined as the average over 2010--2018 and past as the time between the respective beginning of data taking until 1997. We do not find a clear trend. For Esrange \chem{\Delta_{present-past}\left<[O_3]\right>} amounts to $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas to $(-2.5\pm 3.7)\,\unit{ppb}$ \citep[][for comparison]{NILU2003, BER:Ruoho-Airola2015}. The observed trend at Esrange is compatible with our assumed bias correction for the Svanvik ozone climatology.

@@

In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and assess the monthly significance of the ozone concentration anomalies in 2018/19 (Fig.~\ref{fig:anomalies_svanvik}d)). The resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month in Fig~\ref{fig:data_svanvik_2018_2019}. Negative deviations from climatology are shown as a negative percentage. Note that missing bars for certain months are both due to data availability and anomalies not meeting the criteria for significance. The the star indicates resulting anomalies in after the reconstruction of missing values in July. In April/May 2018, ozone was significantly enhanced, while no significant enhancement is found in 2019. In both years, ozone concentrations were significantly lower than average in November.

>>>>

Similarly, we find a skew towards higher \chem{[O_3]} for Svanvik in 2018, while 2019 follows a normal distribution rather well (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}). In conjunction with the meteorological conditions, we therefore can assume that 2019 had been a normal year and that the annual average anomally in 2019 can be used for the bias correction of the Svanvik climatology (\chem{\left<\Delta[O_3]\right>}~=~$(1.2\pm 5.0)\,\unit{ppb}$). To further strengthen this, we look at trends in ozone observations at Esrange and Pallas and compute the average difference between recent years and the mid 1990s (\chem{\Delta_{present-past}\left<[O_3]\right>}). Present is defined as the average over 2010--2018 and past as the time between the respective beginning of data taking until 1997. We do not find a clear trend. For Esrange \chem{\Delta_{present-past}\left<[O_3]\right>} amounts to $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas to $(-2.5\pm 3.7)\,\unit{ppb}$ \citep[][for comparison]{NILU2003, BER:Ruoho-Airola2015}. The observed trend at Esrange is compatible with our assumed bias correction for the Svanvik ozone climatology.

@@

In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and assess the monthly significance of the ozone concentration anomalies in 2018/19 (Fig.~\ref{fig:anomalies_svanvik}d)). The resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month in Fig~\ref{fig:data_svanvik_2018_2019}. Negative deviations from climatology are shown as a negative percentage. Note that missing bars for certain months are both due to data availability and anomalies not meeting the $1\sigma$ criteria. The the star indicates resulting anomalies after the reconstruction of missing data in July. In April/May 2018, ozone was significantly enhanced, while no significant enhancement is found in 2019. In both years, ozone concentrations were significantly lower than average in November. These results mean, that the extensive forest fires in Sweden did not lead to a significant enhancement in \chem{[O_3]} at Svanhovd in 2018.
